---
title: "European Value Study - Survey Response Report"
author: "Jake Martin"
date: "March 30, 2024"
output:
  pdf_document: default
subtitle: Attitudes towards gender roles and immigration
params:
  country: NA
---

```{r setup, global_options, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")

library(haven)
library(tidyverse)
library(kableExtra)
library(texreg)
library(plotly)
library(ggplot2)
library(knitr)
library(sjlabelled)

survey_data <- read_sav('./data/clean_survey_data.sav')

```

## Description

This report summarizes the results of two survey questions concerning attitudes toward immigrants and women with children. The data was collected as part of the European Values Study, which is a "large-scale, cross-national and longitudinal survey research program on how Europeans think about family, work, religion, politics, and society."

This report focuses on two questions that were asked in the survey:

\setlength{\leftskip}{1cm}

***For each of the following statements I read out, can you tell me how strongly you agree or disagree with each.***

*When a mother works for pay, the children suffer*

*When jobs are scarce, employers should give priority to [nationality of respondent] people over immigrants*

***Do you strongly agree, agree, disagree, or strongly disagree?***

\setlength{\leftskip}{0pt}

This report is based on responses from `r nrow(survey_data)` individuals age 18 or older.

For each survey question, "strongly agree" and "agree" responses have been grouped as a binary variable. In some cases, respondents with missing data have been excluded from analysis (e.g. respondents with missing age information are not included in analysis that discusses age).

## Tables
Table 1 summarizes the relationship between education and the percent of respondents who agree that *"When a mother works for pay, the children suffer."* This table indicates that half of people with lower levels of education agree that children suffer when a mother works. 


```{r, echo=TRUE}

survey_data %>% 
  drop_na(education, v72) %>%
  group_by(as_character(education)) %>%
  summarize(mean = round(mean(v72),2),
            count = round(sum(gweight)))%>%
  kbl(col.names = c("Education", "Mean","Weighted n"),
      caption = "Percent agree: Gender roles by Education") %>%
  kable_styling(full_width = FALSE, bootstrap_options = "striped")

```

Table 2  summarizes the relationship between age (grouped into four categories) and the percent of respondents who agree that *"When jobs are scarce, employers should give priority to [nationality of respondent] people over immigrants"* This table indicates that more than 60% of people in each age group agree that employers should give priority to residents over immigrants. 

```{r, results='asis', echo=TRUE}
survey_data %>% 
  drop_na(age, v80) %>%
    mutate(age_4 = case_when(
      (age < 35) ~ "18-34",
      (age > 34 & age < 50) ~ "35-49",
      (age > 49 & age < 65) ~ "50-64",
      (age > 64) ~ "65+")) %>%
  group_by(age_4) %>%
  summarize(mean = round(mean(v80),2),
            count = round(sum(gweight)))%>%
  kbl(col.names = c("Age group", "Mean","Weighted n"),
      caption = "Percent agree: Immigration views by Age") %>%
  kable_styling(full_width = FALSE, bootstrap_options = "striped")
```

## Graphs

Figure 1 displays the relationship with age and the percent of respondents who agree that *"When a mother works for pay, the children suffer."*

```{r, echo=TRUE}

survey_data %>% 
  drop_na(age, v72) %>%
  group_by(age) %>%
  summarize(mean = mean(v72),
            count = sum(gweight)) %>% 
  ggplot(aes(age, mean)) +
           geom_line() + 
  ggtitle("Figure 1: Gender roles by age") 


```

Figure 2 displays the relationship with age and the percent of respondents who agree that *"When jobs are scarce, employers should give priority to [nationality of respondent] people over immigrants."*

```{r, echo=TRUE}

survey_data %>% 
  drop_na(age, v80) %>%
  group_by(age) %>%
  summarize(mean = mean(v80),
            count = sum(gweight)) %>% 
  ggplot(aes(age, mean)) +
           geom_line() + 
  ggtitle("Figure 2: Immigration attitudes by age") 


```

## Regressions

Table 3 illustrates a regression model explaining the relationship between gender roles (Model 1) and immigration (Model 2) by age, age squared, sex, and education.

```{r results='asis', echo=TRUE}

v72_model <- lm(v72 ~ age + age_sq + sex + education_lower + education_higher, data = survey_data)
v80_model <- lm(v80 ~ age + age_sq + sex + education_lower + education_higher, data = survey_data)

v72_model_male_estimate <- summary(v72_model)$coefficients[4,1]
v72_model_male_p <- summary(v72_model)$coefficients[4,4]

v80_model_lowered_estimate <- summary(v80_model)$coefficients[5,1]
v80_model_lowered_p <- summary(v80_model)$coefficients[5,4]

coefnames <- c("Intercept",
               "Age",
               "Age squared",
               "Sex: male",
               "Education: lower",
               "Education: higher")

texreg(list(v72_model,v80_model),
                    caption="Regression model to explain the relationship between views on gender roles and immigration by select demographics",
                    custom.coef.names = coefnames
       )

```

The regression model examining gender roles (Model 1) indicates indicates that there `r ifelse(v72_model_male_p <= .05, "is", "is not")` a statistically significant `r ifelse(v72_model_male_estimate > 0, "positive", "negative")` relationship between men and agreement (Estimate = `r round(v72_model_male_estimate,2)`, p-value = `r round(v72_model_male_p,2)`) that "when a mother works for pay, the children suffer."

The regression model examining immigration attitudes  (Model 2) indicates that there `r ifelse(v80_model_lowered_p <= .05, "is", "is not")` a statistically significant `r ifelse(v80_model_lowered_estimate > 0, "positive", "negative")` relationship between lower education and agreement (Estimate = `r round(v80_model_lowered_estimate,2)`, p-value = `r round(v80_model_lowered_p,2)`) that "When jobs are scarce, employers should give priority to [nationality of respondent] people over immigrants."
